systemfit: A Package for Estimating\\ Systems of Simultaneous Equations in \proglang{R}

### 2007-07-04 paper/trunk/JSS/response_A.tex

%
\subsection{Organize Internals of systemfit}
%   o organize internals of systemfit
% 
%     The internals of a fitted "systemfit" object could be re-organized. Currently,
%     there are a lot of single letter elements - more standard naming as in "lm"
%     objects etc. would be more desirable.
We have renamed many elements of a fitted \code{systemfit} object
to make the naming more standard.
A comparison with the elements returned by \code{lm}
is available in the paper (appendix~A).

%     Also, computing the partial t tests
%     should/could be done in the summary() method, avoiding that certain parameters
%     (not affecting the fit itself, such as degrees of freedom) necessarily have to
%     be specified for model fitting.
Furthermore, we have followed the advice of this referee
and moved the computation of the partial $t$ statistic
from the \code{systemfit()} function to its \code{summary()} method.

% 
%     The technical control parameters could also be collected in a list of arguments, e.g.,
%       control = systemfit.control()
%     that sets suitable defaults but can also easily be changed. If possible `...'
%     arguments could be passed to systemfit.control() assuring backward compatibility
%     and facilitating interactive usage.
Following the suggestion of this referee,
the technical control parameters of \code{systemfit()} are collected
in a single argument \code{control} now.
This argument is a list that contains the technical control parameters,
where suitable default values are set by the function
\code{systemfit.control()}.

%     The names of various
%     arguments should be re-thought, e.g. "tx", "R.restr" and "q.rests" seem rather
%     arbitrary. Similar comments hold for other arguments, see also below for some
%     more comments.
The names of several arguments have been improved.
The arguments to impose parameter restrictions by $R \beta = q$
(former \code{R.restr} and \code{q.restr})
have been renamed to \code{restrict.matrix} and \code{restrict.rhs},
respectively
(following the arguments \code{hypothesis.matrix} and \code{rhs}
of the generic function \code{linear.hypothesis}).
The argument to impose parameter restrictions
by modifying (post-multiplying) the regressor matrix
(former \code{TX}) has been renamed to \code{restrict.regMat}.
Finally, the arguments to specify the methods
for the 3SLS estimation and to calculated the residual covariance matrix
(former \code{formula3sls} and \code{rcovformula})
have been renamed to \code{method3sls} and \code{methodResidCov},
respectively.

@@

% 
%     Having two Grunfeld data sets is confusing as the data used by Theil is simply
%     a subset of the data used by Greene:
%       subset(GrunfeldGreene, firm %in% c("General Electric", "Westinghouse"))
%     Of course, there are different versions of this data set, but neither the paper
%     nor the manual pages of GrunfeldGreene and GrunfeldTheil appropriately discuss
%     the problem. From the manual and the paper, the reader is only able to feel
%     vaguely insecure about the appropriateness of the selected subsets. Also note
%     that a more complete (and slightly different) version of the data set is
%     available in package "Ecdat".
The data set \code{GrunfeldTheil} has been removed,
because it is a subset of the data set \code{GrunfeldGreene}.
We have improved the discussion of different versions of Grunfeld's data set
in the documentation of \code{GrunfeldGreene}
and we have added this discussion in the paper.
Furthermore, we have added a note
that a correct version of this data set with five additional firms
is available as data set \code{Grunfeld} in the \pkg{Ecdat} package.

% 
%
\section{Minor Points and Typos}
We have corrected the paper according to all of these comments.

### 2007-07-12 paper/trunk/systemfit_code.tex
added appendix (section) that shows how to estimate systems of equations with "sem"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@787 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added paragraph that (and how) we use the "Matrix" package for the GLS estimations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@803 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiency of the code}
We have followed \cite{bates04} to make the code faster and more stable.
First, if a formula contains an inverse of a matrix
that is post-multiplied by a vector,
we use \code{solve(A,b)} instead of \code{solve(A) \%*\% b}.
Second, we calculate crossproducts
by \code{crossprod(X)} or \code{crossprod(X,y)}
instead of \code{t(X) \%*\% X} or \code{t(X) \%*\% y},
respectively.

The matrix $\Omega^{-1}$ that is used to compute the estimated
coefficients and their covariance matrix is of size
$( G \cdot T ) \times ( G \cdot T )$
(see sections~\ref{sec:Estimation} and~\ref{sec:Restrictions}).
In case of large data sets, this matrix $\Omega^{-1}$ gets really huge
and needs a lot of memory.
Therefore, we use the following transformation and compute $X' \Omega^{-1}$
by dividing the $X$ matrix into submatrices,
doing some calculations with these submatrices,
adding up some of these submatrices, and
finally putting the submatrices together, so that
\begin{equation}
X' \Omega^{-1}
%= X' \left( \Sigma^{-1} \otimes I \right)
= \sum_{i=1} \left[ \begin{array}{c}
   \sigma^{1i} {X^1} \\
   \sigma^{2i} {X^2} \\
   \vdots \\
   \sigma^{Gi} {X^G} \\
\end{array} \right]' ,
\end{equation}
where $\sigma^{ij}$ are the elements of the matrix $\Sigma^{-1}$,
and $X^i$ is a submatrix of $X$ that contains the rows
that belong to the $i$'s equation.

@@

Function \code{sem} (package \pkg{sem}, \citealp{r-sem-0.9-8})
can be used to estimate structural equation models in \proglang{R}
by Full Information Maximum Likelihood (FIML)
assuming multinormal errors.
A special feature of this function is the estimation
of models with unobserved (°latent°) variables,
which is not possible with \code{systemfit}.
While \code{sem} cannot be used to consistently estimate
systems of simultaneous equations with some endogenous regressors,
it can be used to estimate systems of equations,
where all regressors are exogenous.
However, the latter is rather cumbersome.
Hence, \code{systemfit} is the only function in \proglang{R}
that can be used to estimate systems of simultaneous equations
and it is the most convenient function to estimate
systems of equations with purely exogenous regressors.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiency of the code}
We have followed \cite{bates04} to make the code faster and more stable.
First, if a formula contains an inverse of a matrix
that is post-multiplied by a vector or matrix,
we use \code{solve(A,b)} instead of \code{solve(A) \%*\% b}.
Second, we calculate crossproducts
by \code{crossprod(X)} or \code{crossprod(X,y)}
instead of \code{t(X) \%*\% X} or \code{t(X) \%*\% y},
respectively.

The matrix $\Omega^{-1}$ that is used to compute the estimated
coefficients and their covariance matrix is of size
$( G \cdot T ) \times ( G \cdot T )$
(see sections~\ref{sec:Estimation} and~\ref{sec:Restrictions}).
In case of large data sets, this matrix $\Omega^{-1}$ gets really huge
and needs a lot of memory.
Therefore, we use the following transformation and compute $X' \Omega^{-1}$
by dividing the $X$ matrix into submatrices,
doing some calculations with these submatrices,
adding up some of these submatrices, and
finally putting the submatrices together, so that
\begin{equation}
X' \Omega^{-1}
%= X' \left( \Sigma^{-1} \otimes I \right)
= \sum_{i=1} \left[ \begin{array}{c}
   \sigma^{1i} {X^1} \\
   \sigma^{2i} {X^2} \\
   \vdots \\
   \sigma^{Gi} {X^G} \\
\end{array} \right]' ,
\label{eq:omegaInv}
\end{equation}
where $\sigma^{ij}$ are the elements of the matrix $\Sigma^{-1}$,
and $X^i$ is a submatrix of $X$ that contains the rows
that belong to the $i$'s equation.

Since version 1.0, \pkg{systemfit} uses the \pkg{Matrix} package
\citep{r-matrix-07} by default.
This package provides classes for different types of matrices.
For instance, we choose class \code{dgeMatrix}
(°real matrices in general storage mode°),
for matrices $X_i$ in~(\ref{eq:model-array}),
class \code{dgCMatrix} (°general, numeric, sparse matrices
in the (sorted) compressed sparse column format°)
for matrix $X$ in~(\ref{eq:model-matrices}),
and class \code{dspMatrix}
(°symmetric real matrices in packed storage (one triangle only)°)
for the residual covariance matrix $\SHat$.
If the \pkg{Matrix} package is used,
the possibly huge matrix $\Omega^{-1}$ is no longer a problem,
because it is a sparse matrix
that can be stored in a compressed format (class \code{dgCMatrix}).
Hence, we no longer need the algorithm in equation~(\ref{eq:omegaInv}).
We have tried different ways to calculate a GLS estimator
like in (\ref{eq:sur})
and we found
that the following code is the fastest:
\begin{Schunk}
\begin{Sinput}
> sigmaInv <- solve( residCov )
> xtOmegaInv <- t( xMat ) %*% kronecker( sigmaInv, Diagonal( nObs ) )
> coef <- solve( xtOmegaInv %*% xMat, xtOmegaInv %*%  yVec )
\end{Sinput}
\end{Schunk}
In this code snippet,
\code{residCov} is the residual covariance matrix $\SHat$,
\code{xMat} is the matrix $X$ and
\code{yVec} is the vector $y$ in~(\ref{eq:sur}).

@@

Function \code{sem} (package \pkg{sem}, \citealp{r-sem-0.9-8})
can be used to estimate structural equation models in \proglang{R}
by Full Information Maximum Likelihood (FIML)
assuming multinormal errors.
A special feature of this function is the estimation
of models with unobserved (°latent°) variables,
which is not possible with \code{systemfit}.
While \code{sem} cannot be used to consistently estimate
systems of simultaneous equations with some endogenous regressors,
it can be used to estimate systems of equations,
where all regressors are exogenous.
However, the latter is rather cumbersome
(see appendix~\ref{sec:sem}).
Hence, \code{systemfit} is the only function in \proglang{R}
that can be used to estimate systems of simultaneous equations
and it is the most convenient function to estimate
systems of equations with purely exogenous regressors.
